---
title: "data-analysis-replication"
author: "Sam Smith"
date: "3/2/2022"
output: html_document
---
For this reproducibility project, I will be attempting to replicate the analyses of the following paper:

Kistner, T. M., Zink, K. D., Worthington, S., & Lieberman, D. E. (2021). Geometric morphometric investigation of craniofacial morphological change in domesticated silver foxes. Scientific reports, 11(1), 1-10.

There is an ongoing, long-term, artificial selection study in Russia where researchers have tried to domesticate silver foxes. After only 8-10 generations, they found behavioral differences, and after a few more generations, morphological changes, including differences in craniofacial structures. From this and one other study, researchers have proposed a link between craniofacial changes and domestication. However, Kistner et al. (2021) noted potential issues with this hypothesis including the assumption that the unselected foxes from the original experiment had "wild-type" craniofacial shapes. The authors pointed out that these foxes may have experienced selective pressures from having been farmed for 60+ years. Thus, the goal of their study was to understand the effect of domestication on craniofacial skeletal morphology in silver foxes and specifically identify the magnitude of differences in craniofacial morphology between each of selected, unselected, and wild foxes. They predicted that the differences will be greater between wild and farmed foxes than between unselected and selected groups. They collected 3D landmark data, linear, and endocranial volume measurements on skulls from selected, unselected, and wild foxes. The authors used geometric morphometric analyses on the 3D data to create tangent space coordinates. They then used procrustes ANOVA on these coordinates to find average population differences. Finally, they used a generalized least squares model to analyze the linear and endocranial volume data. They found that the two farmed fox populations were much more similar than either group was to wild foxes, suggesting that unselected fox cranial morphology is not an appropriate proxy for that of wild foxes. They further propose that craniofacial morphology is not a reliable predictor of docility in these foxes.

I will be replicating ... 

The data files have the following naming conventions:

3D landmark coordinates for each specimen : Fox_data_Morphologika.txt

linear measurements for all specimens : Fox_linear_volume_data.csv

Replication data measurements for 3D landmark coordinates: DF_morphologika.txt, DF476_morphologika.txt, DM_morphologika.txt, DM23_morphologika.txt, UF_morphologika.txt, UF1058_morphologika.txt

Replication data measurements for linear measurements: linear_cv_data.csv

Read in 3D landmark coordinate data and look at it. We use the geomorph package because it is a special file type (Morphologika). This was something that I had to google before I understood the file structure. This is presumably familiar to people knowledgeable in this field and wa probably why it was not included in the manuscript.

```{r}
library(geomorph)
f <- "https://raw.githubusercontent.com/samksmith/data-analysis-replication/main/data/Fox_data_Morphologika.txt"
landmark <- read.morphologika(f)
str(landmark)
landmark$coords[,,1]

(n <- length(landmark$labels[,1])) # number of individuals we have 3D landmark data for
(p <- nrow(landmark$coords[,,1])) # number of measured coordinates
(k <- ncol(landmark$coords[,,1])) # number of dimensions to the coordinates
```

This returns a list of 3. The first object in the list has the 3 dimensional coordinates for each sample. The second object has the labels for each sample (id, sex, and group). The final object contains the wireframe values.

Next we need to do a generalized Procrustes analysis. This involves translating landmarks to the origin, rescaling to centroid size, and rotating landmarks using a least squares criterion. this places homologous coordinates in close alignment (in same shape space, centered on the mean shape). From reading the geomorph package manual, this is all done using a single function gpagen.The authors mention in the methods that they use default parameters of the gpagen function.

```{r}
Pro_sup <- gpagen(landmark$coords)
names(Pro_sup)
str(Pro_sup)
```

Next, I need to project these coordinates onto a "linear tangent space." Looks like this is plotting coordinates on principal component axes. From some googling, it looks like there was a function directly for this in early versions of the geomorph package, but it is now depreciated (from v. 3.3.0 on) and they suggest gm.prcomp instead. This function defaults to using an OLS for centering, which I think the authors used (instead of GLS), but I am not 100% sure because they aren't specific. I also print out a summary and plot the amount of variance explained by each PC.

```{r}
tang_space_coords <- gm.prcomp(Pro_sup$coords)
names(tang_space_coords)
summary(tang_space_coords)

# amount of variance explained by different 
pvar <- (tang_space_coords$sdev^2)/(sum(tang_space_coords$sdev^2)) 
names(pvar) <- seq(1:length(pvar))
barplot(pvar, xlab= "Principal Components", ylab = "% Variance")
```

Now, I need to do a permutation-based Procrustes MANOVA. I will regress tangent space coordinates on population identity and sex using the geomorph package. 
Quote below: where to start next time --- 
"To test hypotheses about shape differences among the three populations of foxes, we used a permutation-based Procrustes MANOVA to regress tangent space coordinates on population identity and sex in the geomorph v. 3.3.1 package. Because we are unable to detect significant differences in allometry among populations with a permutation-based Procrustes MANOVA of tangent space coordinates on the inter- action term of population identity and centroid size, the tangent space coordinates were not corrected for any scaling effects (see Supplemental information and Fig S2). Given this result, we control only for isometric size in geometric morphometric analyses (i.e. no correction for scaling in tangent space coordinates) as well as in our linear measurements."


First, I will load in the datasets and do exploratory analyses.

Let's start with the linear measurements for all the specimens. Below, I load in the data as a tibble and show a few lines of raw code.

```{r}
library(tidyverse)
library(dplyr)
library(cowplot)
f <- "https://raw.githubusercontent.com/samksmith/data-analysis-replication/main/data/fox_linear_volume_data.csv"
linear_data <- read_csv(f,col_names = TRUE)
head(linear_data) # what does the raw data look like?
```

There are 9 measurements taken for each skull. We can start by doing some descriptive statistics to get a sense of the central tendency and spread of these linear measurements.

First, let's look at the mean and standard deviations of all the measurements and visualize the central tendency and spread of these linear measurements for each population and sex using boxplots.
```{r}
summary(linear_data[,-c(1:3)]) # show me summary stats for all the skull measurements

p1 <- ggplot(linear_data,aes(x=population,y=zygomatic_width,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Zygomatic width") + geom_point(position=position_jitterdodge(),alpha=0.3)
p2 <- ggplot(linear_data,aes(x=population,y=cranial_vault_width,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Cranial vault width") + geom_point(position=position_jitterdodge(),alpha=0.3)
p3 <- ggplot(linear_data,aes(x=population,y=upper_jaw_width,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Upper jaw width") + geom_point(position=position_jitterdodge(),alpha=0.3)
p4 <- ggplot(linear_data,aes(x=population,y=total_skull_length,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Total skull length") + geom_point(position=position_jitterdodge(),alpha=0.3)
p5 <- ggplot(linear_data,aes(x=population,y=snout_length,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Snout length") + geom_point(position=position_jitterdodge(),alpha=0.3)
p6 <- ggplot(linear_data,aes(x=population,y=cranial_vault_height,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Cranial vault height") + geom_point(position=position_jitterdodge(),alpha=0.3)
p7 <- ggplot(linear_data,aes(x=population,y=endocranial_volume,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Endocranial volume") + geom_point(position=position_jitterdodge(),alpha=0.3)
p8 <- ggplot(linear_data,aes(x=population,y=face_geometric_mean,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Face geometric mean") + geom_point(position=position_jitterdodge(),alpha=0.3)
p9 <- ggplot(linear_data,aes(x=population,y=centroid_size,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Centroid size") + geom_point(position=position_jitterdodge(),alpha=0.3)

plot_grid(p1,p2,p3,p4,p5,p6,p7,p8,p9)

```
Looks like there are sex differences and possible group differences most strongly between the wild population and the other two pops, which seem pretty similar by eye. 

Let's also look to see if there are relationships between these measurements or whether these measurements are independent of each other. It is important to look at this because it will inform the linear modeling later on. 
```{r}
library(corrplot)
cor = cor(linear_data[4:10], use="pairwise.complete.obs")
corrplot.mixed(cor, lower.col = "black",tl.pos='d',number.cex = .7,upper="ellipse")
detach(package:corrplot)
```
All of these variables are correlated, which is unsurprising since they represent difference, non-independent aspect of shape variation measured on the same specimens. This plot looks very similar to what is present in the supplemental figures of the paper.

Before doing modeling these data, the authors size-adjusted the linear and volume measurements because there is an overall difference in isometric size between wild and farmed populations. This was important to do because the authors are interested in effect of docility selection on craniofacial size separate from overall size selection, which is expected to also have an impact on craniofacial size.

Here I will attempt to normalize the 6 linear measurements and cube root of endocranial volume by centroid size (a proxy for overall cranial size). Below, I create a dataframe with the normalized measurements and I slook at the correlation between variables.
```{r}

normal_linear <- tibble("id"=linear_data$id,
                        "population"=linear_data$population,
                        "sex"=linear_data$sex,
                        "zygomatic_width_norm" = linear_data$zygomatic_width / linear_data$centroid_size,
                        "cranial_vault_width_norm" = linear_data$cranial_vault_width / linear_data$centroid_size,
                        "upper_jaw_width_norm" = linear_data$upper_jaw_width / linear_data$centroid_size,
                        "total_skull_length_norm" = linear_data$total_skull_length / linear_data$centroid_size,
                        "snout_length_norm" = linear_data$snout_length / linear_data$centroid_size,
                        "cranial_vault_height_norm" = linear_data$cranial_vault_height / linear_data$centroid_size,
                        "cube_endocranial_volume_norm" = (linear_data$endocranial_volume)^(1/3) / linear_data$centroid_size)
head(normal_linear)
summary(normal_linear[,-c(1:3)]) 

p1 <- ggplot(normal_linear,aes(x=population,y=zygomatic_width_norm,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Normalized zygomatic width") + geom_point(position=position_jitterdodge(),alpha=0.3)
p2 <- ggplot(normal_linear,aes(x=population,y=cranial_vault_width_norm,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Normalized cranial vault width") + geom_point(position=position_jitterdodge(),alpha=0.3)
p3 <- ggplot(normal_linear,aes(x=population,y=upper_jaw_width_norm,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Normalized upper jaw width") + geom_point(position=position_jitterdodge(),alpha=0.3)
p4 <- ggplot(normal_linear,aes(x=population,y=total_skull_length_norm,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Normalized total skull length") + geom_point(position=position_jitterdodge(),alpha=0.3)
p5 <- ggplot(normal_linear,aes(x=population,y=snout_length_norm,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Normalized snout length") + geom_point(position=position_jitterdodge(),alpha=0.3)
p6 <- ggplot(normal_linear,aes(x=population,y=cranial_vault_height_norm,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Normalized cranial vault height") + geom_point(position=position_jitterdodge(),alpha=0.3)
p7 <- ggplot(normal_linear,aes(x=population,y=cube_endocranial_volume_norm,fill=factor(sex))) + geom_boxplot(na.rm=TRUE) + ylab("Normalized, cubed endocranial volume") + geom_point(position=position_jitterdodge(),alpha=0.3)

plot_grid(p1,p2,p3,p4,p5,p6,p7)

library(corrplot)
cor = cor(normal_linear[4:10], use="pairwise.complete.obs")
corrplot.mixed(cor, lower.col = "black",tl.pos='d',number.cex = .7,upper="ellipse")
detach(package:corrplot)
```

My correlation plot looks a lot like the one in the supplementals. INSERT HERE.


The authors then used the normalized data to find whether there were population differences. They used a linear model with a generalized least squares (GLS) estimator from nlme package. They regressed all 7 skull variables simultaneously as correlated responses on population identity and sex. They provide more information in the supplementary methods (Figs S3, S4). 

Because the 7 variable are correlated, it is not appropriate to do separate GLMs for each variable on population identity and sex. The GLS estimator allows correlations between the variables to be accounted for within the model.


```{r}
```

We report estimates of pairwise percent differences between population means for each skull variable. We use this method because the 7 linear and volumetric skull variables were correlated in two ways (see Fig. S3). First, they were measured on the same specimens, and second, they represent non-independent aspects of shape variation. Modeling these response variables in 7 separate general linear models (e.g., ANOVA) would result in biologically unrealistic inferences because these correlations would be artificially fixed at zero. In addition, since skull variables exhibited varying amounts of dispersion, the GLS model allowed for different residual variances for each response variable.